---
title: "Initial Korean covid 19"
author: "Genesis Alfaro"
date: "06/12/24"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
options(digits=2)
require(tidyverse)
require(tigerstats)
require(lubridate)
kRdata <- read.csv(file = 'kRdata.csv')
kRdata <- kRdata %>%
mutate(across(where(is.character), as.factor))

```
### Data is from the Patient_Info dataset from South Korea 1 Jan to 1 May 2020
### This data was retrieved from a downloadable zip file from kaggle

Some of the data was uncoded for sex - sex must be just male or female
77 of the data entries were coded sex  as blank, so we omit these.
state of a patient must be died or alive - we adjust that as well   
  
# Your Question - Is dying independent of sex?
As i am looking at this data, the appropriate variables to use would be either sex all lower cased or Sex with an uppercase S and in my opinion I belive the Sex with the uppercase S because the variable are just "f" and "M" when the other one it spells out the whole word "female" and "male". The second variable that would apply to this would be the result because the only nes you could use are alive or dead, only 2 options.





## Graphical results - use ggplot commands to do this  
```{r}
dd2 <- kRdata %>% group_by(result,Sex) %>% summarize(count=n()) %>% mutate(prcnt=count/sum(count))


# the group_by followed by summarize(count=n())

basicC <- ggplot(dd2,aes(x=result,y=count,fill=Sex))
basicC + geom_bar(stat="identity",position="dodge")

#Now for percentage plot

basicCC <- ggplot(dd2,aes(x=result,y=prcnt*100,fill=Sex)) 
basicCC + geom_bar(stat="identity", position = "dodge")
```

As we look at the graphs, we see that there were more women than men in the alive graph. The dead pile shows more men were dead than females. The alive pile represents approximately 55% of people that were female and 45% that were male. When looking at the dead graph it shows approximately 37% that were female while approximately 63% were male, even though there were more females that caught covid, the ones dying were the males 2-1.
  





  
## Numerical results


## Inferential results 




## Conclusions

